---
title: "Chemical Data Analysis & Plot"
author: "Marisa"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    theme: flatly
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # shows code
                      warning = FALSE, # excludes warnings
                      message = FALSE, # excludes messages
                      fig.path = "/..output/")
```

## Introduction
In this document, I will:  

1. Show what libraries I use in my R code for data analysis  
2. Describe how to read in data & average it in R  
3. Show my analyzed data in a table and a plot  

## Load Libraries
These are the libraries we will be using in our code today.

If you don't have them, you can install them using **install.packages("** _[packagename]_ **")**

```{r}
# Load Libraries
library(here)
library(tidyverse)
library(kableExtra)

```
## Getting our data
Today we will be using the Chem Data from Week 3.
```{r}
# Reads in Chem data .csv file and assigns it as ChemData
ChemData <- read_csv(here("Week_06", "data", "chemicaldata_maunalua.csv"))

# Lets us look at some of our data
glimpse(ChemData)
```

## Analyzing our data!
```{r, include = FALSE}

ChemData <- ChemData %>% 
  
  # excludes NA values
  drop_na() %>% 
  
  # separates Tide_time into individual Tide and Time columns
  separate(col = Tide_time,
           into = c("Tide","Time"),
           sep = "_", 
           remove = TRUE) %>% 
  
  # Selects only the columns for Site, Season, Time, and Salinity
  select(Site, Season, Time, Salinity) %>% 
  
  view(ChemData)
  
```

I have gone ahead and selected for _only_ the parameters we are interested in:

1. Site  

2. Season (Fall/Spring)  

3. Time (Day/Night)  

4. Salinity  


This is the data we will work with today.

#### Averaging Data

Let's first **average** the Salinity data:
```{r}
ChemData <- ChemData %>% 
# Groups by site, season, and time
  group_by(Site, Season, Time) %>% 
  # calculates averages in Salinity, removing NA values
  summarize(mean_Sal = mean(Salinity, na.rm = TRUE))
```


Great!
Below is a table that shows our data, which includes our parameters of interest and our freshly-calculated averages:

```{r, echo = FALSE}
ChemData %>% 
  kbl() %>% # makes kable table
  kable_classic_2() %>%
  # specifies style and position of table
  kable_styling(full_width = FALSE, 
                position = "center",
                htmltable_class = "lightable-striped")
  
```
### Plot

Now we can **plot** our data. I have gone ahead and created a **bar plot** using the data we just averaged:

```{r, Sal_plot, out.width="75%", fig.align='center', fig.cap="This plot shows average salinity at different sites (BP, W) during different seasons (Fall, Spring) at different times of day (day, night).", echo = FALSE}
ChemData_plot <- ggplot(data = ChemData,
                        mapping = aes(x = Site,
                                      y = mean_Sal,
                                      fill = Time))+
  
  geom_bar(position = "dodge", stat = "identity")+ # Makes bar plot
  
  # Below subsets plot by Season
  facet_wrap(~Season)+
  
  # Below creates labels for plot elements
  labs(x = "Site",
       y = "Salinity",
       fill = "Time of Day",
       title = "Average Salinity at different Times of Day during Fall vs Spring Season")+
  
  # specifies fill for Day vs Night
  scale_fill_manual(values = c("darkorange1", "darkmagenta"))+
  
  # Below specifies color and size for various plot elements
  theme(axis.title = element_text(size = 12, color = "saddlebrown"),
        legend.title = element_text(size = 10, color = "saddlebrown"),
        panel.background = element_rect(fill = "seashell"),
        strip.background = element_rect(fill = "khaki2"))


ChemData_plot # displays plot
```

Ta-da!
We now have our plot.
 
 
 
 
 

**_Thanks for reading!_**

