---
title: "R Markdown with Penguins Part 2"
author: "Marisa"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    theme: "sandstone"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      # start from Week_06 folder
                      # it will go back one flder (..), then go into                             output
                      fig.path = "../output/")
```

### General Info
This document will be used to practice working with figures and other elements in RMarkdown.

### Load Libraries
Use this code to load the libraries you will need today:
```{r}
# Load Libraries
library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
```

### Let's load & look at our data

```{r}
glimpse(penguins)
```

### Create a plot!
<!-- Put name in {r } because chunk is associated with an output (plot).
Can do {,out.width="50%"} to change width of figure as a percent
Can do {, fig.width=5, fig.height=7} to change height and width of figure
Can do {, fig.asp=1.2} to change aspect ratio
Can do {, fig.align='center'} to center the plot-->

```{r penguin_plot, out.width="50%", fig.align='center', fig.cap="this is my figure", }
plot <- ggplot(data = penguins,
       mapping = aes(y = flipper_length_mm,
                     x = island,
                     fill = species))+
  geom_bar(position = "dodge", stat = "identity")+
  
  labs(y = "Flipper length (mm)",
       x = "Island",
       fill = "Species")

plot
```
## Making Tables
You can make tables by hand using |, :, and -

| Time  | Session | Topic |
|:------|:-------:|------:|
| left  | center  | right |
| 01:00 | 1       | Anatomy|
| 01:50 |         | Break  |
| 2:00  | 2       | Tables |

You can also make Tables with kable
_Note: you must install the package "kableExtra"_

```{r}
penguins %>% # the data we are using
  # calculates mean of flipper legnth grouped by species
  group_by(species) %>% 
  summarize(flipper_mean = mean(flipper_length_mm, na.rm = TRUE)) %>%
  kbl() %>% # makes a table
  kable_classic() %>% 
  # changes table elements (colors, text, etc.)
  row_spec(2, bold = T, color = "gray92", background = "cornflowerblue") %>% 
  # Makes width of the table NOT the width of the page
  kable_styling(full_width = FALSE)
```

