---
title: 'Homework: Two Useful Functions'
author: "Marisa Mackie"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

### Introduction
In this document I will be showing two useful functions.

* Function 1: Summary Statistics of Data Grouped by 2 Variables

* Function 2: Creating Bar Plot

### Libraries
```{r}
# Load libraries
library(tidyverse)
library(here)
library(palmerpenguins)
```

### Data
We will be using the palmerpenguins dataset.
```{r}
glimpse(penguins)
```


### Function #1: Statistics when Grouped by 2 Variables  
#### df_stats2()

With this function, you can create a new dataframe showing summary statistics of your data, such that you are grouping by only 2 variables.

*Note: This function will drop missing (NA) values*

The arguments of this function include:

* data = the data you want to use

* gv1 & gv2 = the 2 variables you would like to group by ("grouping variables")

* x = the variable you would like to calculate summary statistics for


The output is a new dataframe that will show the count, average, variance, and standard deviation of your desired variable, grouped by the 2 grouping variables.

```{r, df_stats2}
# Variables: gv1 and gv2 are the two variables you wish to group by
# x is the variable you would like to calculate summary statistics for

df_stats2 <- function(data, gv1, gv2, x) { # specifies fxn name & args
  
  df <- data %>% # grabs from data
    
  drop_na() %>% # drops NA values
    
  group_by({{gv1}},{{gv2}}) %>% # groups by 2 variables
    
  summarize(n = nrow({{x}}), # count
            avg = mean({{x}}, na.rm = TRUE), # averages data
            variance = var({{x}}, na.rm = TRUE), # calculates variance
            std_dev = sd({{x}}, na.rm = TRUE)) # calculates standard deviation
  
  return(df) # returns dataframe
    
}
```

Let's test it with our penguin data:

I would like to group by (1) species and (2) sex, and I want to calculate summary statistics for flipper length.

```{r, include = TRUE}
# Create new dataframe "peng_stats" using df_stats2 function to our penguin data.
peng_stats <- df_stats2(data = penguins, gv1 = species, gv2 = sex, x = flipper_length_mm)

glimpse(peng_stats)

```


### Function #2: Creating Bar Plot  
#### barplot()

With this function, you can plot your data as a basic bar plot.
```{r, barplot function}
barplot <- function(data, x, y, fill, plot_title, x_axis, y_axis, legend_title){ # specifies fxn name & args
  
  plot <- data %>% 
    
    # generates plot & specifies variables based on input
    ggplot(data, mapping = aes(x = {{x}}, 
                               y = {{y}}, 
                               fill = {{fill}}))+
    
    geom_bar(stat = "identity", position = "dodge")+ # specifies bar plot
    
    # specifies size & color of plot elements
    theme(plot.title = element_text(size = 15, face = "bold"),
          axis.title = element_text(size = 12, color = "saddlebrown"),
          legend.title = element_text(size = 10, color = "saddlebrown"),
          panel.background = element_rect(fill = "lightcyan"))+
    
    # specifies plot labels based on input
    labs(title = plot_title,
         x = x_axis,
         y = y_axis,
         fill = legend_title)
  
  return(plot) # returns plot
}
```

Let's test it with the peng_stats data output from Function 1.

I will be using this function to plot the average flipper length for each species, between sexes. I will also specify the titles of the plot, axes, and legend that I want.
```{r}

# Uses barplot function to plot desired data
barplot(data = peng_stats, x = species, y = avg, fill = sex, plot_title = "Average Flipper Length for Penguin Species", x_axis = "Species", y_axis = "Average Flipper Length (mm)", legend_title = "Sex")

# saves plot to output folder
ggsave(here("Week_09","output","ExamplePlotForFunction2.png"))

```

###### Thanks for reading!