---
title: 'Homework: For Loops & Purrr'
author: "Marisa Mackie"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    toc: TRUE
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## Introduction
In this document I will:

* Bring in all 4 files

* Calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool. 

* Use both a for loop and map() functions


## Load Libraries
```{r}
# Load libraries
library(tidyverse)
library(here)
library(janitor)
```


## For loop approach:

I will first do each of these steps individually, and then I will construct my For loop.

#### Read in the data for 1 file

First, let's read in one of the data files and take a look at it.

```{r}
# read in first data file
TP1 <- read_csv(here("Week_12","data","Homework","TP1.csv"))

glimpse(TP1)
```

#### Clean the names

Uh oh. It looks like one of our variable names have special characters like periods "." in them, which can make it harder to work with. Let's change those periods "." to underscores "_"

An easy way to do this is to use the **clean_names()** function in the **Janitor** package.

```{r}
TP1_clean <- TP1 %>% 
  clean_names() # cleans column names by converting them to snake_Case

glimpse(TP1_clean)
```
#### For Loop:

Let's first read in all the data files.
```{r}
pool_path <-here ("Week_12", "data", "homework")

# lists all the files in the path that contain ".csv"
TPfiles <- dir(path = pool_path, pattern = ".csv")
TPfiles
```


Let's pre-allocate space so that our For loop can store the output somewhere.

```{r}
# creates empty dataframe where number of rows = number of files, and has 3 columns
pool_data <- data.frame(matrix(nrow = length(TPfiles), ncol = 5))

# gives column names to the pool_data dataframe
colnames(pool_data)<-c("filename","mean_temp","stdev_temp","mean_light","stdev_light")

pool_data
```


Now that we have this, let's create our For loop to do this for all the files!

```{r}
for(i in 1:length(TPfiles)){ # loops from 1st to last file in pool_data
  # reads in files, pastes path with / to file name
  file_data <- read_csv(paste0(pool_path, "/", TPfiles[i])) %>% 
    clean_names() # cleans the names, converts to snake_case
  
  # gets the filename based on the current file
  pool_data$filename[i] <- TPfiles[i]
  
  # calculates mean temp for current file
  pool_data$mean_temp[i] <- mean(file_data$temp_c, na.rm = TRUE)
  
  # calculates mean light intensity for current file
  pool_data$mean_light[i] <- mean(file_data$intensity_lux, na.rm = TRUE)
  
  # calculates std deviation for temp for current file
  pool_data$stdev_temp[i] <- sd(file_data$temp_c, na.rm = TRUE)
  
  # calculates std deviation for light intensity for current file
  pool_data$stdev_light[i] <- sd(file_data$intensity_lux, na.rm = TRUE)
}

pool_data
```


## Map functions approach:

Now we will do the same thing using map functions!

```{r}
# sets path
pool_path <- here("Week_12", "data", "homework")

# lists all the files in the path that contain ".csv"
TPfiles <- dir(path = pool_path, pattern = ".csv", full.names = TRUE)

# calculates mean & std deviation of temperature & light for each file
TP_summary <- TPfiles %>%
  
  set_names() %>% # set the id of each list to the filename
  
  map_df(read_csv,.id = "filename") %>% # maps it to a dataframe and puts ID in column "filename"
  
  clean_names() %>% 
  
  group_by(filename) %>% # groups by filename
  
  # calculates mean of temp & light intensity, dropping NAs
  summarize(mean_temp = mean(temp_c, na.rm = TRUE),
            stdev_temp = sd(temp_c, na.rm = TRUE),
            mean_light = mean(intensity_lux, na.rm = TRUE),
            stdev_light = sd(intensity_lux, na.rm = TRUE)) %>% 
  
  # renames the filenames to only include the last 7 characters, i.e., TPn.csv
  mutate(filename = str_sub(filename, start = -7))


TP_summary
```

## Check for accuracy

Let's compare them to make sure they are correct:

```{r}
glimpse(pool_data)
glimpse(TP_summary)
```
Looks good!


###### Thanks for reading!
