---
title: 'Homework: Map of Cats on the Move!'
author: "Marisa"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    toc: TRUE
    theme: architect
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../output/")
```

### Introduction
This is my homework assignment for week 7. I will be creating a map using the **_Cats on the Move!_** dataset.

### Libraries
We will need to load the following libraries:
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggmap)
library(ggsn)
```
```{r, echo = FALSE}
# Using API to access Google Maps data
register_google(key = "your key here", write = TRUE)
```

### Data
We first need to load in our Cat data
```{r}
# loads our cat data
cats <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')

# lets us look at our cats data
view(cats)
```

Since this cat data was taken from the United Kingdom (UK), let's select map data for just the UK. Specifically, our cat data is mostly in the counties of Cornwall & Devon.

```{r}
# gets longitude & latitude data from the UK, specifically Cornwall county
geocode("Cornwall")

# Creates dataframe with the lon & lat found previously
Cornwall <- data.frame(lon = -5.052713, lat = 50.26605)
```

### Analysis
Next, let's just select for the relevant cat data:  

 1. tag ID (identifier for cat)
 2. event ID
 3. location (longitude)
 4. location (latitude)
 
```{r}
# Selects for the following columns in our cats data
cats <- cats %>% 
  select(tag_id, event_id, location_long, location_lat)

view(cats)
```

I want to plot 10 random cats. To do this, I will randomly generate 10 numbers and select from the list of cat IDs. Then I will use the associated longitude & latitude data.
```{r}
# Arranges cat tag IDs alphabetically and lists distinct cat IDs
  cats_unique <- cats %>% 
  distinct(tag_id, .keep_all = FALSE) %>% 
  arrange(tag_id)

# Selects 10 random cat IDs from the alphabetized cat ID list, without repeats
  cats_random <- cats_unique[sample(1:101, 10, replace = FALSE),1]

view(cats_random)

# Joins randomly selected cat ID list with associated longitude & latitude data values into new dataset
ten_cats <- left_join(cats_random, cats) %>% 
  group_by (tag_id, event_id)

view(ten_cats)
```


### Plot
Now, let's plot our points on our map!

```{r, CatMap, fig.align = "center",fig.cap = "This map shows movement of 10 randomly selected cats in the United Kingdom, specifically near Cornwall county"}
# Gets map of Cornwall county, zoomed out to see neighboring counties
Cornwall_map <- get_map(Cornwall, zoom = 9, maptype = "watercolor")

CatMap <- ggmap(Cornwall_map)+
  
  # Plots cat data as a scatterplot
  geom_line(data = ten_cats,
             # x & y are coordinates of cat location
             aes(x = location_long,
                 y = location_lat,
                 # color = tag id for each cat
                 color = tag_id),
             size = 0.8)+
  
  # sets colors for points manually
  scale_color_manual(values = c("darkmagenta","midnightblue","tomato4","mediumblue","violetred4","green4","chocolate3","purple3","turquoise4","seagreen4"))+
  
  # sets labels for plot, axes, etc.
  labs(x = "Longitude",
       y = "Latitude",
       color = "Cat Name (Tag ID)",
       title = "Movement of 10 Random Cats in Cornwall county, United Kingdom")+
  
  # sets size and color for plot elements
  theme(axis.title = element_text(size = 10, color = "saddlebrown"),
        plot.title = element_text(size = 15, color = "saddlebrown"))

CatMap
```

###### Data Source:

Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563

McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732

###### Map Source: ggmap

D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2.
  The R Journal, 5(1), 144-161. URL
  http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf



###### Thanks for reading!